# Audio features — Spotify

Spotify has developed a number of metrics that assess the sonic characteristics of any given songs. These features are described in their [API documentation](https://developer.spotify.com/documentation/web-api/reference/tracks/get-audio-features/). The R package [`spotifyr`](https://github.com/charlie86/spotifyr) provides a great interface to access these features.

In order to use spotifyr, you must have API credentials. See the [README](https://github.com/charlie86/spotifyr) for more instructions. Below I reference an `.Renviron`file I created that hosts my Spotify API credentials.

```{r}
library(spotifyr)

access_token <- get_spotify_access_token(
  client_id = Sys.getenv("SPOTIFY_CLIENT_ID"),
  client_secret = Sys.getenv("SPOTIFY_CLIENT_SECRET")
)
```


One current limitation of the API is that it requires a song's identification number to retrieve audio features. In an effort to simplify the process of retrieving the audio features for each song, I created a function that we can use with `map()` inside of a `mutate()` call. Note that this is a very useful pattern commit to memory. 


```{r}
track_audio_features <- function(artist, title, type = "track") {
  search_results <- search_spotify(paste(artist, title), type = type)
  
  track_audio_feats <- get_track_audio_features(search_results$id[[1]]) %>%
    dplyr::select(-id, -uri, -track_href, -analysis_url)
  
  return(track_audio_feats)
}
```

Let's take a moment to undersand what is happening in this function. First, we take the `artist` and `title` arguments and create a search string to provide to `search_spotify()`. This returns a data frame object for the top search results. Next, we take the track ID from the first search result (this works under the assumption that the spotify_search is robust enough to identify the track based on the artist and song name), and is fed to the `get_track_audio_features()` function.

Let's see how this works on it's own by search spotify for a single song. Below I look for the song _Nocturne: Lost Faith_ by the metalcore band _Invent, Animate_.

```{r}
ia <- search_spotify("Invent, Animate Nocturne: Lost Faith", type = "track")
ia
```

Now we can take the ID column and feed it to `get_track_audio_features()`. Note that there will often be more than one search result, hence the subsetting within the function. 

```{r}
get_track_audio_features(ia$id)
```

How does the self-defined function perform?

```{r}
track_audio_features("Invent, Animate", "Nocturne: Lost Faith")
```

What happens if a song is not found?

```{r eval=FALSE}
track_audio_features("definiely not", "in spotify")
```

We are returned a very uninformative error. As we will iterate through our data frame of 600 songs, there is a good chance that at least one track will not be found in Spotify. To cope with this we can use the `possibly()` function from `purrr` which allows functions to fail gracefully. By wrapping our function in `possibly()`, we can specify what is returned in the case of an error.

```{r}
possible_feats <- possibly(track_audio_features, otherwise = tibble())
```

Here we define another function `possible_feats()` which, when it encounters an error, will return an empty tibble. The empty tibble is important as we will later have to unnest the data frame.

Now we are prepared to iterate through all of the tracks and fetch the audio features. Note this is not a quick operation creating a great opportunity to pour yourself another cup of coffee.

```{r eval=FALSE}
chart_analysis <- charts %>%
  mutate(audio_features = map2(artist, title, possible_feats)) %>%
  unnest() %>% 
  as_tibble()
```

```{r include=FALSE}
chart_analysis <- read_rds("data/chart_analysis.rds")
```

```{r}
glimpse(chart_analysis)
```

